The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth

### 2021-09-30 paper/paper-jsdse/paper-jsdse.tex
corrected the wages description

The employment data comprises of three variables, i.e., \texttt{total\ hours\ of\ work\ per\ week}, \texttt{number\ of\ jobs\ that\ an\ individual\ has}, and \texttt{mean\ hourly\ wage}.

\texttt{hours\ worked\ per\ week} initially has only one version per job, no choice from 1979 to 1987 (QES-52A). From 1988 onward, when we had more options, we chose the variable for total hours, including time spent working from home (QES-52D). However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead. In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we use these.

The same way was also deployed to tidy the rate of wage by year and by ID. The difference is that the hourly rate has only one version of each year. The hours of work and the hourly rate are then joined to calculate the number of jobs that a respondent has and their mean hourly wage. Some observations have 0 in their hourly rate and work hours, which is considered an invalid value. Thus, these observations are set to be NA. Besides, there are some observations with unusual hours of work. Here, we also censored the observations with \textgreater{} 84 hours of work a week to be NA.

Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculated based on the availability of the \texttt{rate\_per\_hour} information. For example, the number of jobs of ID 1, based on \texttt{hours\_work}, is 2. However, since the information of hourly rate of \texttt{job\_02} is not available, the number of jobs is considered as 1.

Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight. However, there is a lot of missing value in \texttt{hours\_work} variable. In that case, we only calculated the mean hourly wage based on the arithmetic/regular mean method. Hence, we created a new variable to flag whether the mean hourly wage is weighted or regular. Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as arithmetic mean.

@@

The employment and demographic variables are then joined. We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called \texttt{wages}.

@@

{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/yowie/paper/figures/flowchart}

@@

\caption{The stages of data cleaning from the raw data to get three datasets contained \texttt{[CENSORED]}. ``\# of individuals'' means the number of respondents included in each stage, while ``\# of observations'' means the number of rows in the data. The color represents the stage of data cleaning in statistical value chain (M. P. J. van der Loo and de Jonge 2021). Pink, blue, and green represents the raw, input, and valid data, respectively.}\label{fig:flow-chart}
\end{figure}

>>>>

Our target variables for the employment are to obtain respondent's mean hourly wage (\texttt{mean\_hourly\_wage}), the number of jobs (\texttt{number\_of\_jobs}) and the total hours of work per week (\texttt{total\_hours}) for each survey year. As the data only reports up to 5 jobs for each respondent, the maximum number of jobs is capped at 5.

From 1979 to 1987, \texttt{new\_data\_qnames} only contain one version of hours worked per week for each job (in the variables with names starting with \texttt{QES-52A}). From 1988 onward, we selected the total hours worked per week, including hours working from home (\texttt{QES-52D}). However, in 1993, this variable was missing for the first and last job so we selected to use \texttt{QES-52A} instead. In addition, 2008 only had jobs 1-4 for the \texttt{QES-52D} variable, so we use only these.

The hourly wages are in the variables beginning with \texttt{HRP} in \texttt{new\_data\_qnames}. As a respondent may have multiple jobs, the \texttt{mean\_hourly\_wage} is computed as a weighted average of the hourly wage for each job with the number of hours worked for each job as weights (provided that the information on number of hours is available); if number of hours worked for any job is missing, then the \texttt{mean\_hourly\_wage} is computed as a simple average of all available hourly wage. Prior to computing the mean hourly wage, we undertook a number of steps to treat unusual observations as described below:

\begin{itemize}
\tightlist
\item
  If the hourly rate is recorded as 0, we set wage as missing.
\item
  If the total hours of worked for the corresponding job is greater than 84 hours, we set the wage and hour worked as missing.
\end{itemize}

The number of jobs (\texttt{number\_of\_jobs}) for each respondent per year is computed from the number of non-missing values of hourly wage \emph{and} hours worked. If either the hourly wage or hours worked is missing, we do not tally this.

@@

The employment and demographic variables are then joined. This data are further filtered to the cohort who completed the education up to 12th grade and participated at least five rounds in the survey. We save the resultant wage data on this cohort as \texttt{wages}.

@@

{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/yowie/paper/figures/flowchart_blind}

@@

\caption{The stages of data cleaning from the raw data to get three datasets contained \texttt{[CENSORED]}. ``\# of individuals'' means the number of respondents included in each stage, while ``\# of observations'' means the number of rows in the data. The color represents the stage of data cleaning in statistical value chain (M. P. J. van der Loo and de Jonge 2021). Pink, blue, and green represents the raw, input, and valid data, respectively.}\label{fig:flow-chart-blind}
\end{figure}

### 2021-09-30 paper/paper.Rmd
Add conditional chunk for blinded version.
corrected the wages description

The employment data comprises of three variables, i.e., `total hours of work per week`, `number of jobs that an individual has`, and `mean hourly wage`.

`hours worked per week` initially has only one version per job, no choice from 1979 to 1987 (QES-52A). From 1988 onward, when we had more options, we chose the variable for total hours, including time spent working from home (QES-52D). However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead. In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we use these.

@@

The same way was also deployed to tidy the rate of wage by year and by ID. The difference is that the hourly rate has only one version of each year. The hours of work and the hourly rate are then joined to calculate the number of jobs that a respondent has and their mean hourly wage. Some observations have 0 in their hourly rate and work hours, which is considered an invalid value. Thus, these observations are set to be NA. Besides, there are some observations with unusual hours of work. Here, we also censored the observations with > 84 hours of work a week to be NA.

@@

Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculated based on the availability of the `rate_per_hour` information. For example, the number of jobs of ID 1, based on `hours_work`, is 2. However, since the information of hourly rate of `job_02` is not available, the number of jobs is considered as 1.

Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight. However, there is a lot of missing value in `hours_work` variable. In that case, we only calculated the mean hourly wage based on the arithmetic/regular mean method. Hence, we created a new variable to flag whether the mean hourly wage is weighted or regular. Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as arithmetic mean.

@@

The employment and demographic variables are then joined. We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called `wages`.

@@

```{r flow-chart, fig.cap="(ref:flow-chart)", out.width="85%"}
```

>>>>

Our target variables for the employment are to obtain respondent's mean hourly wage (`mean_hourly_wage`), the number of jobs (`number_of_jobs`) and the total hours of work per week (`total_hours`) for each survey year. As the data only reports up to 5 jobs for each respondent, the maximum number of jobs is capped at 5.

From 1979 to 1987, `new_data_qnames` only contain one version of hours worked per week for each job (in the variables with names starting with `QES-52A`). From 1988 onward, we selected the total hours worked per week, including hours working from home (`QES-52D`). However, in 1993, this variable was missing for the first and last job so we selected to use `QES-52A` instead. In addition, 2008 only had jobs 1-4 for the `QES-52D` variable, so we use only these.

@@

The hourly wages are in the variables beginning with `HRP` in `new_data_qnames`. As a respondent may have multiple jobs, the `mean_hourly_wage` is computed as a weighted average of the hourly wage for each job with the number of hours worked for each job as weights (provided that the information on number of hours is available); if number of hours worked for any job is missing, then the `mean_hourly_wage` is computed as a simple average of all available hourly wage. Prior to computing the mean hourly wage, we undertook a number of steps to treat unusual observations as described below:

* If the hourly rate is recorded as 0, we set wage as missing.
* If the total hours of worked for the corresponding job is greater than 84 hours, we set the wage and hour worked as missing.

The number of jobs (`number_of_jobs`) for each respondent per year is computed from the number of non-missing values of hourly wage *and* hours worked. If either the hourly wage or hours worked is missing, we do not tally this.

@@

@@

The employment and demographic variables are then joined. This data are further filtered to the cohort who completed the education up to 12th grade and participated at least five rounds in the survey. We save the resultant wage data on this cohort as `wages`.

@@

```{r flow-chart-blind, fig.cap="(ref:flow-chart)", out.width="85%", eval = CENSOR}
```

### 2021-09-30 paper/paper.tex
corrected the wages description

The NLSY79 data contains a large scope of variables but for our aim, the scope required is limited to demographic profiles and wages data. More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described in \newline\Figure \ref{fig:source-nav}.

>>>>

The NLSY79 data contains a large scope of variables but for our aim, the scope required is limited to demographic profiles and wages data. More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described in Figure \ref{fig:source-nav}.
